Ms 1

### 2017-11-01 ms1/ms1_display.Rmd
refactoring display

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
import::from(tools, md5sum)
library(forcats)
```

# Abstract

```{r abstract, cache.extra=md5sum('tables/model_abx.tsv')}
pctize = function(df, denom_name) {
  for (n_name in Filter(function(x) str_detect(x, '^n_'), names(df))) {
    pct_name = str_replace(n_name, '^n_', 'pct_')
    df[[pct_name]] = df[[n_name]] / df[[denom_name]] * 100
  }
  df
}

totals = read_tsv('tables/totals.tsv') %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west)

claims_by_abx = read_tsv('tables/claims_by_abx.tsv')

est2pct = function(x) (exp(x) - 1)*100

mult_est = function(df, m) {
  mutate(df, estimate=m*estimate, std.error=sqrt(m)*std.error)
}

model_ci = function(df) {
  df %>%
    mutate(pct_change=est2pct(estimate),
           ci=1.96*std.error,
           ci_low=est2pct(estimate - ci),
           ci_high=est2pct(estimate + ci),
           change_str=sprintf('%.2f (%.2f -- %.2f)', pct_change, ci_low, ci_high))
}

claims_overall = claims_by_abx %>%
  select(-n_bene) %>%
  group_by(year) %>%
  summarize_at(vars(n, cpkp), sum) %>%
  mutate(antibiotic='overall')

bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', 'azithromycin', 'levofloxacin'), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round(100*(`2014`-`2011`)/`2011`, 2)) %>%
  kable
```

# Beneficiary characteristics

```{r unique_bene_pde}
round_millions = function(x) round(x / 1e6, 2)
pn = function(x) prettyNum(x, big.mark=',')

n_unique_bene = scan('tables/n_unique_bene.tsv')

claims_by_fill = read_tsv('tables/claims_by_fill.tsv')
n_pde = sum(claims_by_fill$n)
n_refills = claims_by_fill %>% filter(!first_fill) %$% n
```

There were `r round_millions(n_unique_bene)` million unique beneficiaries. They had `r round_millions(n_pde)` million PDEs, of which `r round_millions(n_refills)` million were refills.

A star indicates that the trend (*t*-test for mean difference between years equal to zero) is significant after multiple hypothesis correction. A dot indicates significance ($p < 0.05$) before multiple hypothesis correction. Mean age decreased during the study.

```{r bene_char_table}
trend_p = function(x) tidy(t.test(diff(x)))$p.value
totals_trends = totals %>%
  gather('var', 'val', -year) %>%
  arrange(var, year, val) %>%
  group_by(var) %>%
  summarize(p=trend_p(val)) %>%
  mutate(adj.p=p.adjust(p, method='BH')) %>%
  mutate(sig=case_when(.$adj.p < 0.05 ~ '*',
                       .$p < 0.05 ~ '.',
                       TRUE ~ ''))

totals %>%
  pctize('n_bene') %>%
  select(year, n_bene, mean_age, sd_age, mean_n_cc, sd_n_cc,
         pct_female, pct_white, pct_dual,
         pct_region_south, pct_region_midwest, pct_region_northeast, pct_region_west) %>%
  mutate_at(vars(starts_with('pct_')), function(x) round(x, digits=1)) %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=','),
         mean_age=round(mean_age, digits=1)) %>%
  mutate_at(vars(mean_n_cc, sd_age, sd_n_cc),
            function(x) round(x, digits=2)) %>% 
  gather('var', 'val', -year) %>% spread(year, val) %>% # transpose
  left_join(totals_trends %>% select(var, sig), by='var') %>%
  kable
```

# Patterns of consumption

All consumption rates are reported as claims per thousand people per year (CPKP).

## Claims by abx

The top 10 antibiotics account for most of the claims:

```{r top_10_fraction}
claims_by_abx %>%
  group_by(antibiotic) %>%
  summarize(n=sum(n)) %>%
  arrange(desc(n)) %>%
  mutate(f=n/sum(n), cumf=cumsum(f)) %>%
  head(10) %>%
  kable
```

```{r claims_by_abx}
top_abx = claims_by_abx %>%
  group_by(antibiotic) %>%
  summarize(cpkp=mean(cpkp)) %>%
  arrange(desc(cpkp)) %$%
  head(antibiotic, 10)

claims_by_top_abx = claims_by_abx %>%
  bind_rows(claims_overall) %>%
  mutate(antibiotic=fct_relevel(fct_other(antibiotic, keep=c('overall', top_abx)),
                                c('overall', top_abx, 'Other'))) %>%
  group_by(year, antibiotic) %>%
  summarize_at(vars(n, cpkp), sum) %>% ungroup()
```

```{r claims_by_top_abx_figs}
claims_by_top_abx %>%
  filter(antibiotic=='overall') %>%
  ggplot(aes(x=year, y=cpkp)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(975, 1525), breaks=c(1000, 1125, 1250, 1375, 1500)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, all abx')

claims_by_top_abx %>%
  filter(antibiotic %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'trimethoprim/sulfamethoxazole', 'levofloxacin')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, top 5')

claims_by_top_abx %>%
  filter(antibiotic %in% c('amoxicillin/clavulanate', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
  ggplot(aes(x=year, y=cpkp, color=antibiotic)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
  #ylim(0, 140) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, next 5')
```

@@

```{r claims_by_demography}
change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  filter(term=='year') %>%
  select(name, adj_change=change_str)

claims_by = function(x) {
  read_tsv(str_interp("tables/claims_by_${x}.tsv")) %>%
    rename_(.dots=list(name=x)) %>%
    select(name, year, cpkp) %>%
    filter(year %in% c(2011, 2014)) %>%
    spread(year, cpkp) %>%
    mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
    mutate_at(vars(`2011`, `2014`), round) %>%
    mutate(group=x)
}

# do claims by age group separate
# this could be baked into the analysis step
claims_by_age = read_tsv('tables/claims_by_age.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  mutate(name=case_when(between(.$age, 65, 75) ~ 'age65_75',
                        between(.$age, 76, 85) ~ 'age76_85',
                        between(.$age, 86, 95) ~ 'age86_95',
                        .$age >= 96 ~ 'age96_')) %>%
  group_by(year, name) %>%
  summarize_at(vars(n_bene, n_claims), sum) %>%
  ungroup() %>%
  mutate(cpkp=1000*n_claims/n_bene) %>%
  select(year, name, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
  mutate_at(vars(`2011`, `2014`), round) %>%
  mutate(group='age')

# use a common interface for all the others
bind_rows(claims_by_age,
          claims_by('sex'),
          claims_by('race'),
          claims_by('region')) %>%
  left_join(change_demog, by='name') %>%
  select(group, name, `2011`, `2014`, pct_change, adj_change) %>%
  kable
```

That last table could also use an "overall" row:

```{r}
claims_overall %>%
  filter(year %in% c(2011, 2014)) %>%
  mutate(cpkp=round(cpkp)) %>%
  kable
```

# Adjusted consumption changes by drug

Adjusted changes were computed using a Poisson regression predicting claims per capita. The unit of analysis were individual beneficiaries in each year. The covariates in the regression were age, number of chronic conditions, dual eligibility, sex, race (white or not), and region.

## By drug

### Trends in appropriateness

Just by eye, it looks like the changes in appropriateness are small.

```{r tier_graph}
pde_tier_table = dx_pde_counts %>%
  filter(has_dx_em) %>%

pde_tier_table = read_tsv('raw_data/dx_pde_counts.tsv') %>%
  group_by(year,

pde_tier_table = read_tsv('tables/pde_approp_table.tsv') %>%
  group_by(year, antibiotic) %>%
  mutate(f=n/sum(n)) %>%
  ungroup()

pde_tier_table %>%
  filter(antibiotic %in% top_abx) %>%
  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
  geom_bar(stat='identity') +
  facet_grid(~antibiotic) +
  theme_classic()
```

```{r approp_trends}
raw_adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error)

# fraction inappropriate in 2011
f_i = pde_tier_table %>%
  filter(year==2011) %>%
  mutate(inapp=tier==3) %>%
  group_by(inapp) %>%
  summarize(n=sum(n)) %>%
  mutate(f=n/sum(n)) %>%
  filter(inapp) %$%
  f

# fold change in consumption
pde_trend = raw_adjusted_changes %>%
  filter(antibiotic=='overall') %>%
  # multiplications by 3 done above
  mutate(ci=1.96*std.error,
         rr=exp(estimate),
         rr_lo=exp(estimate-ci),
         rr_hi=exp(estimate+ci))

risk_ratio_f = function(odr, p) odr / (1 + p * (odr - 1))

pde_inapprop_trend = read_tsv('tables/pde_inapprop_trend.tsv') %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error,
         ci=1.96*std.error,
         or=exp(estimate),
         or_lo=exp(estimate-ci),
         or_hi=exp(estimate+ci))

d_f = risk_ratio_f(pde_inapprop_trend$or, f_i)
d_f_lo = risk_ratio_f(pde_inapprop_trend$or_lo, f_i)
d_f_hi = risk_ratio_f(pde_inapprop_trend$or_hi, f_i)

inapprop_changes = data_frame(
  antibiotic=c('overall_inapp', 'overall_app'),
  pct_change=c(d_c * d_f, (1.0 - d_f * f_i) / (1.0 - f_i) * d_c),
  ci_low=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
  ci_high=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
) %>%
  mutate_if(is.numeric, rr2pct)
```

@@

Bars show adjusted trends in usage, including refills. The dots shows unadjusted trends.

```{r adjusted_changes}
adjusted_changes = raw_adjusted_changes %>%
  model_ci() %>%
  arrange(desc(pct_change)) %>%
  select(antibiotic, pct_change, ci_low, ci_high)

adjusted_change_levels = adjusted_changes %>%
  filter(antibiotic != 'overall') %$%
  { c('overall', 'overall_app', 'overall_inapp', antibiotic) }

unadjusted_changes = bind_rows(claims_by_abx, claims_overall) %>%
  filter(antibiotic %in% c('overall', top_abx), year %in% c(2011, 2014)) %>%
  select(antibiotic, year, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(change=`2014`-`2011`,
         fold_change=change/`2011`,
         pct_change=100*fold_change) %>%
  select(antibiotic, pct_change)

adjusted_changes %>%
  bind_rows(inapprop_changes) %>%
  mutate(antibiotic=as.character(antibiotic)) %>%
  mutate(antibiotic=factor(antibiotic, levels=rev(adjusted_change_levels))) %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_bar(stat='identity', fill='white', color='black') +
  geom_errorbar(width=0.5) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in consumption 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

The bars show the adjusted change in consumption.

### Adjusted changes in usage table

```{r adjusted_changes_table}
adjusted_changes %>%
  bind_rows(inapprop_changes) %>%
  mutate(antibiotic=factor(antibiotic, levels=adjusted_change_levels)) %>%
  arrange(antibiotic) %>%
  kable

adjusted_az_decrease = adjusted_changes %>%
  filter(antibiotic=='azithromycin') %>%
  pull(pct_change) %>%
  round(2)

adjusted_levo_increase = adjusted_changes %>%
  filter(antibiotic=='levofloxacin') %>%
  pull(pct_change) %>%
  round(2)
```

Consumption of azithromycin fell by `r adjusted_az_decrease`% over the entire study period. Levofloxacin consumption increased by `r adjusted_levo_increase`% over the entire study period.

>>>>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE,
                      error=FALSE, autodep=TRUE)
import::from(knitr, kable)
```

# Population characteristics

```{r pop_chars}
pop_chars = read_tsv('tables/pop_chars.tsv') %>%
  select(-sem_age, -sem_n_cc) %>%
  mutate_at(vars(n_female:n_region_midwest), function(x) x / .$n_bene * 100)

trend_p = function(x) tidy(t.test(diff(x)))$p.value

pop_char_trends = pop_chars %>%
  select(year, n_bene, mean_age, mean_n_cc, n_female:n_region_midwest) %>%
  gather('key', 'value', -year) %>%
  arrange(key, year) %>%
  group_by(key) %>%
  summarize(p=trend_p(value)) %>%
  mutate(adj.p=p.adjust(p, method='BH'))

pop_chars %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=',')) %>%
  mutate_at(vars(mean_age, sd_age, n_female:n_region_midwest), function(x) format(x, digits=1, nsmall=1)) %>%
  mutate_at(vars(mean_n_cc, sd_n_cc), function(x) format(x, digits=2, nsmall=2)) %>%
  gather('key', 'value', -year) %>%
  spread(year, value) %>%
  mutate(key=factor(key, levels=names(pop_chars))) %>%
  arrange(key) %>%
  kable(caption='Population characteristics')

pop_char_trends %>%
  kable(caption='Trends in population characteristics')
```

There are `r scan('tables/n_unique_bene.txt') %>% prettyNum(big.mark=',')` unique beneficiaries.

**swo**: Double-check that number.

There are **swo: PDES** and first fills.

# Trends in drug usage

## Raw trends

```{r raw_drug_trends}
claims_by_abx = read_tsv('tables/claims_by_abx.tsv') %>%
  mutate(abx=fct_inorder(factor(abx)))

abx_levels = levels(claims_by_abx$abx)

claims_by_abx %>%
  filter(abx=='overall') %>%
  ggplot(aes(x=year, y=cpkp)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(975, 1525), breaks=c(1000, 1125, 1250, 1375, 1500)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, all abx')

claims_by_abx %>%
  filter(abx %in% c('azithromycin', 'ciprofloxacin', 'amoxicillin', 'cephalexin', 'tmpsmx', 'levofloxacin')) %>%
  ggplot(aes(x=year, y=cpkp, color=abx)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 250), breaks=c(0, 50, 100, 150, 200, 250)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, top 5')

claims_by_abx %>%
  filter(abx %in% c('amoxclav', 'doxycycline', 'nitrofurantoin', 'clindamycin')) %>%
  ggplot(aes(x=year, y=cpkp, color=abx)) +
  geom_point() + geom_line() +
  scale_y_continuous(limits=c(0, 100), breaks=c(0, 50, 100)) +
  theme_minimal() +
  ggtitle('Unadjusted consumption, next 5')
```

## Trends in appropriateness

```{r approp_trends}
# fold change in consumption
pde_trend = drug_trends %>%
  filter(abx=='overall') %>%
  mutate(rr=exp(estimate),
         rr_lo=exp(ci_lo),
         rr_hi=exp(ci_hi))

# fraction inappropriate in 2011
risk_ratio_f = function(or, p) or / (1 + p * (or - 1))
f_i = scan('tables/inapprop_risk_2011.txt')

inapprop_trend = read_tsv('tables/inapprop_trend.tsv') %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  mutate(ci=1.96*std.error,
         ci_lo=estimate-ci,
         ci_hi=estimate+ci) %>%
  mutate(or   =exp(estimate),
         or_lo=exp(ci_lo),
         or_hi=exp(ci_hi)) %>%
  mutate(rr   =risk_ratio_f(or, f_i),
         rr_lo=risk_ratio_f(or_lo, f_i),
         rr_hi=risk_ratio_f(or_hi, f_i))

d_f = inapprop_trend$rr
d_f_lo = inapprop_trend$rr_lo
d_f_hi = inapprop_trend$rr_hi

inapprop_trends = data_frame(
  abx=c('overall_inapp', 'overall_app'),
  y   =c(d_c    * d_f,    (1.0 - d_f    * f_i) / (1.0 - f_i) * d_c),
  ymin=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
  ymax=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
) %>%
  mutate_if(is.numeric, rr2pct)

kable(inapprop_changes)
```

## Adjusted trends

```{r drug_trends}
est2pct = function(x) (exp(x) - 1)*100

drug_trends = read_tsv('tables/trend_models.tsv') %>%
  filter(term=='year') %>%
  arrange(estimate) %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  mutate(ci=1.96*std.error,
         ci_lo=estimate-ci,
         ci_hi=estimate+ci) %>%
  mutate(y=est2pct(estimate),
         ymin=est2pct(ci_lo),
         ymax=est2pct(ci_hi)) %>%
  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))

drug_trends %>%
  select(abx, y, ymin, ymax) %>%
  mutate_if(is.numeric, function(x) format(x, digits=2, nsmall=2)) %>%
  kable(caption='Drug trends: Percent change from 2011 to 2014')

abx_trends_levels = drug_trends %>%
  filter(abx != 'overall') %$%
  { c('overall', 'overall_app', 'overall_inapp', rev(abx)) }

drug_and_inapprop_trends = drug_trends %>%
  select(abx, y, ymin, ymax) %>%
  bind_rows(inapprop_trends) %>%
  mutate(abx=factor(abx, levels=rev(abx_trends_levels)))

drug_and_inapprop_trends %>%
  ggplot(aes(x=abx, y=y, ymin=ymin, ymax=ymax)) +
  geom_bar(stat='identity', fill='white', color='black') +
  geom_errorbar(width=0.5) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in consumption 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))

drug_and_inapprop_trends %>%
  arrange(desc(abx)) %>%
  kable(caption='Percent changes; same data as in figure')
```

# Supplement

@@

```{r claims_by_demography}
claims_by = function(by) {
  read_tsv(str_interp("tables/claims_by_${by}.tsv")) %>%
    mutate(group_type=by) %>%
    rename(group=!!rlang::sym(by)) %>%
    filter(year %in% c(2011, 2014)) %>%
    select(group_type, group, year, cpkp) %>%
    spread(year, cpkp) %>%
    mutate_at(vars(`2011`, `2014`), round)
}

trends_by_demography = read_tsv('tables/demography_models.tsv') %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  mutate(ci=1.96*std.error,
         ci_lo=estimate-ci,
         ci_hi=estimate+ci) %>%
  mutate(y=est2pct(estimate),
         ymin=est2pct(ci_lo),
         ymax=est2pct(ci_hi)) %>%
  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
  select(group_type, group, change)

# overall trend row
overall_change = drug_trends %>% filter(abx=='overall') %>% pull(change)
overall_row = claims_by_abx %>%
  filter(abx=='overall', year %in% c(2011, 2014)) %>%
  mutate(cpkp=round(cpkp)) %>%
  spread(year, cpkp) %>%
  mutate(group_type='overall', group='overall', change=overall_change) %>%
  select(group_type, group, `2011`, `2014`, change)

bind_rows(
  claims_by('age_group'),
  claims_by('sex'),
  claims_by('race'),
  claims_by('region')
) %>%
  left_join(trends_by_demography, by=c('group', 'group_type')) %>%
  bind_rows(overall_row, .) %>%
  kable
```

# ***Earlier content

@@
